---
title: "p8105_hw6_zz3167"
author: "Zhiyi Zhu"
date: "2023-11-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

#### Import data

```{r}
# Download the Central Park weather data
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated from these data:

* r̂2
* log(β̂1∗β̂2)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂2 and log(β̂1∗β̂2). Note: `broom::glance()` is helpful for extracting r̂2 from a fitted regression, and `broom::tidy()` (with some additional wrangling) should help in computing log(β̂1∗β̂2).

#### Use bootstrap samples to  produce estimates 

```{r}
boot_strap_df = 
  weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df))) 
 

r_squared_df = 
  boot_strap_df |>
  mutate(results = map(models, broom::glance)) |>
  select(-strap, -models) |>
  unnest(results)

log_beta_df = 
  boot_strap_df |>
  mutate(results = map(models, broom::tidy)) |>
  select(-strap, -models) |>
  unnest(results) |>
  select(.id, term, estimate) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |>
  mutate(log_beta_pro = log(tmin * prcp))

r_squared_result = 
  r_squared_df |>
  summarize(r_squared_estimate = mean(r.squared))

log_b1b2_result = 
  log_beta_df|>
  summarize(log_b1b2_estimate = mean(log_beta_pro, na.rm = TRUE))

r_squared_result |>
  knitr::kable()

log_b1b2_result |>
  knitr::kable()
```

* Comment:For log(β1*β2), there are 3329 non-finite values, and calculation of parameter statistics is performed on the non-NA values.

#### Plot the distribution of your estimates

```{r}
# Distribution of r squared estimates
r_squared_df |>
  ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(
    title = "Distribution of Estimated R Squared" ,x = "R Squared Estimates")
```

* Describe the plot: The distribution is approximately normal distribution and left-skewed. We observe that r^2 values are close to 0.92, indicating a strong linear relationship. This suggests that both 'tmin' and 'prcp' serve as reliable indicators of 'tmax'.

```{r}
# Distribution of log(b1b2) estimates
log_beta_df |>
  na.omit() |>
  ggplot(aes(x = log_beta_pro)) +
  geom_density() +
  labs(title = "Distribution of Estimated log(beta1*beta2)", x = "log(beta1*beta2)")
```

* Describe the plot: The distribution is concentrated near -5 but has a heavy tail extending to low values, which may be related to the outliers included in the bootstrap sample.

#### Provide a 95% confidence interval 

```{r}
r_squared_df |>
  summarize(
    r2_CI_lower = quantile(r.squared, 0.025),
    r2_CI_upper = quantile(r.squared, 0.975)
  ) |>
  knitr::kable()
```

```{r}
log_beta_df |>
  na.omit() |>
  summarize(
    log_b1b2_CI_lower = quantile(log_beta_pro, 0.025),
    log_b1b2_CI_upper = quantile(log_beta_pro, 0.975)) |>
  knitr::kable()
```

### Problem 3

#### Load and clean the data

```{r}
#load and clean data
birth_weight = 
  read_csv("data/birthweight.csv") |>
  janitor::clean_names() |> 
  mutate(
    babysex = case_match(
      babysex,
      1 ~ 'male',
      2 ~ 'female'),
    frace = case_match(
      frace, 
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other", 
      9 ~ "Unknown"),
    malform = case_match(
      malform, 
      0 ~ 'absent',
      1 ~ 'present'),
    mrace = case_match(
      mrace, 
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other"))

# check NAs
sum(is.na(birth_weight))
```

There’s no missing values in this dataset.

#### Propose a regression model for birthweight

```{r}
full_model = 
  lm(bwt ~ ., data = birth_weight) 

full_model |>
  broom::tidy() |>
  knitr::kable() 

stepwise_model = 
  full_model |>
  MASS::stepAIC(direction = "backward", trace = FALSE)

stepwise_model |>
  broom::tidy() |>
  knitr::kable()
```

* Describe modeling process: First, I fit a linear model with all the other variables as predictors. Then, using `stepwiseAIC` function to perform stepwise model selection based on the full model. From the result table, the selected predictors are: `babysex`, `bhead`, `blength`, `delwt`, `fincome`, `gaweeks`, `mheight`, `mrace`, `parity`, `ppwt`, `smoken`.

#### Show a plot of model residuals against fitted values 

```{r}
birth_weight |> 
    add_predictions(stepwise_model) |> 
    add_residuals(stepwise_model) |> 
    ggplot(aes(x = pred, y = resid)) +
    geom_point()  + 
    geom_smooth(method = "lm",se = FALSE) + 
    labs(title = "Residuals versus Fitted Values", 
       x = "Predictions", 
       y = "Residuals")
```

#### Compare your model to two others:

* One using length at birth and gestational age as predictors (main effects only)
* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
model_1 = lm(bwt ~ blength + gaweeks, data = birth_weight)

model_1 |>
  broom::tidy() |>
  knitr::kable()
```

```{r}
model_2 = lm(bwt ~ bhead * blength * babysex, data = birth_weight)

model_2 |>
  broom::tidy() |>
  knitr::kable()
```

* Make this comparison in terms of the cross-validated prediction error

```{r}
cv_df =
  crossv_mc(birth_weight, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

compare_df = 
  cv_df |>
  mutate(
    model_0  = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + mheight + mrace + parity + ppwt + smoken, data = .x)),
    model_1  = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
    model_2  = map(train, ~lm(bwt ~ bhead*blength*babysex, data = .x))) |>
  mutate(
    rmse_0 = map2_dbl(model_0, test, ~rmse(model = .x, data = .y)),
    rmse_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
    rmse_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y))) 

compare_df |>
  summarize(
    mean_rmse_0 = mean(rmse_0),
    mean_rmse_1 = mean(rmse_1),
    mean_rmse_2  = mean(rmse_2)
  ) |>
  knitr::kable()
```

```{r}
compare_df |>
  select(starts_with("rmse")) |>
  mutate(
    model_0 = rmse_0,
    model_1 = rmse_1,
    model_2 = rmse_2
  ) |>
  pivot_longer(
    model_0:model_2,
    names_to = "model",
    values_to = "rmse"
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin(aes(fill = model)) +
  labs(
    title = "RMSE of 3 Different Models"
  )
```

* Comment: From the plot, we could conclude that model_0 (which construct using stepwiseAIC) has the smallest RMSE. The model_1 (using `gaweeks` and `blength` as predictors) has the largest RMSE. It indicates that model_0 is the most reliable one to fit.




